Estimating the Financial Impact of Wildfire Smoke on US National parks

### 2020-05-18 Markdown Documents/Manuscript.Rmd
moving things around

```{r echo=FALSE, cache=TRUE}
head(FEdat[c(1009,256,10,800,655),])
```

@@

We operationalized this model in rstanarm
```{r eval=FALSE}
feefit<-stan_glm(SingleVisitorEntryFee~Year+Park,family="Poisson",data=d,chains=3,iter=6000)
```

We then used this to predict what the fees would have been for every year we didn't have
```{r cache=TRUE}
fillData<-FEdat[,1:2]

FEdat$PredictedFee<-round(apply(posterior_predict(feefit,fillData),2,mean),digits = 0) #assuming this is for a new column

@@

Our predicted fees correlate well with our observed fees that we have
```{r echo = FALSE, cache=TRUE}
xx<-na.omit(FEdat)
 mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                  plot.title = element_text( size=18, color="black",face="bold"),
                  axis.title.x = element_text( size=18),
                  axis.title.y = element_text( size=18),
                  axis.text=(element_text(color="black", size=14)),
                  legend.title = element_text(colour="black", size=18),
                  legend.text = element_text( size = 14))
xx%>%
  ggplot(.,aes(x=SingleVisitorEntryFee,y=round(PredictedFee),digits=0))+
  geom_jitter(size=2, alpha=0.3)+
  geom_smooth(method="lm", se=FALSE, colour="red")+
  geom_text(aes(x=5,y=15),size=10,label=paste("cor =",round(cor(round(xx$PredictedFee,digits=0),xx$SingleVisitorEntryFee),digits=2)))+
  ggtitle("Predicted vs Observed Single Visitor Entry Fee")+theme_classic()+mytheme+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_y_continuous(name="Predicted Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))+
  scale_x_continuous(name="Observed Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))
```

@@

```{r warning=FALSE, cache=TRUE}
fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
Data$trendvis<-predict(fits,date=Data$date)
Data$VisDiff<-Data$RecreationVisits-Data$trendvis
dat<-Data
```

@@

This was also operationalized in `rstanarm`
```{r eval=FALSE, cache=TRUE}
fit<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=dat,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)

@@

### 2.2.2. Smoke, Fee, & Visitation Projections

@@

For all significant effects (90% C.I. doesn't overlap 0), we see a negative effect of smoke on the residual visitation to National Parks.

@@

We can then go back and apply these estimates effects to the observed visitation given a minimum level of smoke observed for each park to estimate what the visitation may have been given optimal air quality.

@@

#get the visitation observed and predicted yearly totals (same scale as fee data)
VEstDat<-dat%>%group_by(UnitCode,Year)%>%
  summarise(RecVisits=sum(RecreationVisits),PredVisits = sum(PredNoSmoke))%>%
  filter(UnitCode %in% fitsig$Par)

@@

summaryFinancialLoss<-summaryFinancialLoss[,c(6,4,5)]
summaryFinancialLoss$IncomeDif<-currency(summaryFinancialLoss$IncomeDif)
names(summaryFinancialLoss)<-(c("National Park","Income Loss Estimate (1980-2018)", "Percent Income Loss (1980-2018)"))
summaryFinancialLoss$`Percent Income Loss (1980-2018)`<-paste0(summaryFinancialLoss$`Percent Income Loss (1980-2018)`,'%')
formattable::formattable(summaryFinancialLoss)
```

>>>>

@@

We operationalized this model in rstanarm

We then used this to predict what the fees would have been for every year we didn't have
```{r cache=TRUE, include=FALSE}
fillData<-FEdat[,1:2]

FEdat$PredictedFee<-apply(posterior_predict(feefit,fillData),2,mean) #assuming this is for a new column

@@

Our predicted fees correlate well with our observed fees that we have
```{r echo = FALSE, cache=TRUE}
xx<-na.omit(FEdat)
 mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                  plot.title = element_text( size=18, color="black",face="bold"),
                  axis.title.x = element_text( size=18),
                  axis.title.y = element_text( size=18),
                  axis.text=(element_text(color="black", size=14)),
                  legend.title = element_text(colour="black", size=18),
                  legend.text = element_text( size = 14))
xx%>%
  ggplot(.,aes(x=SingleVisitorEntryFee,y=PredictedFee))+
  geom_jitter(size=2, alpha=0.3)+
  geom_smooth(method="lm", se=FALSE, colour="red")+
  geom_text(aes(x=5,y=15),size=10,label=paste("cor =",round(cor(xx$PredictedFee,xx$SingleVisitorEntryFee),digits=2)))+
  ggtitle("Predicted vs Observed Single Visitor Entry Fee")+theme_classic()+mytheme+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_y_continuous(name="Predicted Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))+
  scale_x_continuous(name="Observed Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))
```

@@

```{r warning=FALSE, include=FALSE, cache=TRUE}
fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
Data$trendvis<-predict(fits,date=Data$date)
Data$VisDiff<-Data$RecreationVisits-Data$trendvis
dat<-Data
```

@@

This was also operationalized in `rstanarm`
```{r eval=FALSE, cache=TRUE,include=FALSE}
fit<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=dat,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)

@@

### 2.2.2. Model Validation

### 2.2.3. Smoke, Fee, & Visitation Projections

@@

## 3.1. Parameter Estimates
For all significant effects (90% C.I. doesn't overlap 0), we see a negative effect of smoke on the residual visitation to National Parks.

@@

## 3.2. Backcasting
We can then go back and apply these estimates effects to the observed visitation given a minimum level of smoke observed for each park to estimate what the visitation may have been given optimal air quality.

@@

#get the visitation observed and predicted yearly totals (same scale as fee data)
VEstDat<-dat%>%group_by(UnitCode,Year)%>%
  summarise(RecVisits=sum(RecreationVisits),PredVisits = sum(PredNoSmoke50CI))%>%
  filter(UnitCode %in% fitsig$Par)

@@

summaryFinancialLoss<-summaryFinancialLoss[,c(6,4,5)]
summaryFinancialLoss$IncomeDif<-currency(summaryFinancialLoss$IncomeDif)
names(summaryFinancialLoss)<-(c("National Park","Income Loss Estimate (1980-2018)", "Percent Income Loss (1980-2018)"))
summaryFinancialLoss$`Percent Income Loss (1980-2018)`<-paste0(summaryFinancialLoss$`Percent Income Loss (1980-2018)`,'%')
formattable::formattable(summaryFinancialLoss)
```
## 3.3. Model Validation

## 3.4. Future Projections
